A Computational Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-03-30 paper/paper.tex
modify in response to XXX

\begin{itemize}
  \item Both the predictive coding account and POMDPs
  are process theories backed up by plausible
  neurobiological evidence that
  view the brain as a ``statistical organ''
  generalizing from the past to new experiences.
  \item Both provide a parsimonious explanation why the
  human brain decreases processing load devoted to incoming information
  when the environment is predictable and allocate increased
  attentional resources when novel stimuli are encountered.
  Similarly, both propose explanations why
  predicting the future is inherently
  linked to information from the past.
  \item Sensory experience is a generative process from both views.
  In predictive coding, sensory perception of the external world
  is a generative experience due to modulatory top down experience at
  various hierarchical levels of sensory input processing.
  In our RL view, the (partially observed)
  environmental model is incorporated in the POMDP,
  \textit{which can be fully recovered based on the last
  state alone}.
  \item Both naturally expose a mechanisms of brain plasticity in that
  neuronal wiring gets increasingly adapted
  when faced by unprecedented environmental
  challenges.
  \item The hierarchical aspects from predictive coding
  is re-expressed in POMDPs in form of
  nested prediction of probable upcoming actions and rewards.
  \item Both model the consequences of action. In RL, the horizon of that
  look into the future is explicitly manifested in the $\gamma$ parameter
  in the Bellman equation, but not
  explicitly modeled in the PC account that
  tends to emphasize prediction about the
  immediate future.
  \item Mismatch negativity of PC is immanent in reward prediction error
  by the difference between actually predicted reward of an action given
  a state and the reward predicted by non-linear regression with graded
  discounting of future rewards.
  \item Adapting the neuronal connections for improved top-down modulation
  takes the concrete form of gradient computation
  \textit{and back-propagation} in MDPs and RL,
  although this may not be neurobiologically plausible \textbf{XXX (danilo) what is not plausible. Be specific.}
  \item In sum,
  the MDP account may serve as a concrete conceptualization of the PC account
  in cognitive neuroscience. MPDs have the advantage of exposing an explicit
  mechanism of the horizon of future considerations or
  how the internal model of the world gets updated,
  and collapses sensory input processing and action output preparation.
  %
  As such, our account can be viewed as a RL implemention of the
  prediction coding hypothesis. \textbf{XXX (danilo) Reformulate. Also in general, you should be clear whether
    you're talking about MDPs, POMDPS, both, or neither. These are being mixed up a lot.}
\end{itemize}

@@

\begin{itemize}
  \item Both the semantic hypothesis and MDPs expose mechanisms of
  how alternative decision trees could be mentally explored.
  \item In both semantic hypothesis and MDPs,
  there is no evidence to indicate that predictions of various
  levels of complexity, abstraction, timescale and purpose
  use mechanisms that are qualitatively different. This concurs with
  DMN activity increases across time, space, and content domains
  demonstrated in many neuroimaging studies. The semantic hypothesis
  and RL account provide explanations why hippocampus lesion does
  not only impair retrieving memories, but also hypothetical and future
  thinking \citep{hassabis2007patients}.
  \item The notion of semantic or knowledge association is
  incorporated into the MDP as the Markov property,
  that is, the current state directly results from the
  agent's history of states and actions. The learned
  value matrix and action transition schedules drive
  stimulus processing and action choice in the present.
  \item While mental scene construction implies a distinction between
    environmental perception and internally generated mind-wandering\footnote{
      \textbf{XXX (danilo) This is more than just a footnote. Make this part of the paper
      and factorize the lot. Also you probably need a citation here!!!}
  ``A computational model of how such a process might be structured
  is far from being defined, but it will probably require a form of
  regulation by which perception of the current world is suppressed
  while simulation of possible alternatives are constructed,
  followed by a return to perception of the present.''\citep{buckner2007self}},
  the MDPs naturally integrate the former egocentric
  (more related to current action, state and reward) and the later
  allocentric (more related to past and future actions, states, and rewards)
  angles on the world in a same optimization problem.
  \item The semantic account of DMN function does not offer
  a mechanistic explanation how explicit world knowledge and semantic analogies thereof
  lead to prediction of future actions and states.
  \item In contrast to existing accounts on semantics and
  mental scene construction, the random and creative aspects of DMN function
  are explained in MDPs by the advantages of stochastic optimization.
  \item The semantic hypothesis does not explain why memory recall
  for scene construction in humans is typically fragmentary and noisy
  instead of accurate and reliable.
\textbf{XXX (danilo) Is this a loose opinion or a fact ? What do you mean by ``probably better models of the world'' ?}  Yet, the MDP framework provides an algorithmic explanation in that
  stochasticity of the parameter space search implemented
  by Monte Carlo solvers provably yields better models of the world.
  That is, the purposeful stochasticity of policy and value estimation
  in MDPs provides a candidate explanation for why humans
  have evolved imperfect noisy memories
  as the more advantageous adaptation.
  \item While both semantic and MDP account propose memory-based internally
  generated information for probabilistic mental models of action outcomes,
  only MDPs render explicit the grounds on which the final action is
  eventually chosen (i.e., the estimated cumulative reward).
  \item In sum, episodic scene construction according to the semantic
  hypothesis is lacking an explicit time and incentive model;
  neither does it explain how randomness of human mental experience
  can be beneficial.
\end{itemize}

@@

\textbf{XXX (danilo) Your focus in this subsection ought to be on the SC itself not POMPDs, etc.}
 \begin{itemize}
    \item Processing social information has been proposed to underlie the
    physiological baseline of human brain function
    \citep{schilbach2008minds}. This was later challenged by observing
    analogues of the DMN in monkeys \citep{mantini2011default}
    and rats \citep{lu2012rat}, with
    supposedly less advanced social capacities.
    The MDPs parsimoneously explain the dominance of social content in
    human mental scenes by their extremely high relevance to humans
    reflected by high values for social information in the value matrix
    \citep{baker2009action}.
    \item How can a same neurobiological circuit be equally important
    for baseline house-keeping functions and specific task performance?
    Our computational account can explain why the DMN is implicated
    in both a goal-directed task and an idlying rest cognitive set,
    if environmental relevance is processed, manipulated, and retained
    also as a baseline function of human mental activity.
    In tasks, the policy and value matrices are updated to optimize short-term action,
    whereas, at rest, these parameter updates may
    improve especially mid- and long-term action.
    \item The sentinel account does not provide a formal mechanisms of
    how attention and memory resources are exactly reallocated when
    encountering a salient environmental stimulus. \textbf{XXX remove the next sentence. Danilo: can you help me to correct the sentence instead.
    There appears to be a cool link between cognition and formal theory
    here.}
    In contrast,
    the Bellman recursion is weighted at each step by
    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
    are recursively explored by the agent and which decision trees are neglected.
    This directly implies allocation of attention and computational load.
    \item In sum,
    MDPs imply a ``radar'' function of monitoring the environment for salient information
    in general and explain social thinking as a physiological brain baseline
    as an important special case, without excluding other topics
    contemplated in continuous mental activity.
 \end{itemize}

>>>>

\begin{itemize}
  \item Both the predictive coding account and POMDPs
  are process theories backed up by plausible
  neurobiological evidence that
  view the brain as a ``statistical organ''
  generalizing from the past to new experiences.
  \item Both provide a parsimonious explanation why the
  human brain decreases processing load devoted to incoming information
  when the environment is predictable and allocate increased
  attentional resources when novel stimuli are encountered.
  Similarly, both propose explanations why
  predicting the future is inherently
  linked to information from the past.
  \item Sensory experience is a generative process from both views.
  In predictive coding, sensory perception of the external world
  is a generative experience due to modulatory top down experience at
  various hierarchical levels of sensory input processing.
  In our RL view, the (partially observed)
  environmental model is incorporated in the POMDP,
  \textit{which can be fully recovered based on the last
  state alone}.
  \item Both naturally expose a mechanisms of brain plasticity in that
  neuronal wiring gets increasingly adapted
  when faced by unprecedented environmental
  challenges.
  \item The hierarchical aspects from predictive coding
  is re-expressed in POMDPs in form of
  nested prediction of probable upcoming actions and rewards.
  \item Both model the consequences of action. In RL, the horizon of that
  look into the future is explicitly manifested in the $\gamma$ parameter
  in the Bellman equation, but not
  explicitly modeled in the PC account that
  tends to emphasize prediction about the
  immediate future.
  \item Mismatch negativity of PC is immanent in reward prediction error
  by the difference between actually predicted reward of an action given
  a state and the reward predicted by non-linear regression with graded
  discounting of future rewards.
  \item Adapting the neuronal connections for improved top-down modulation
  takes the concrete form of gradient computation
  \textit{and back-propagation} in MDPs and RL,
  although the neurobiological plausibility of
  the back-propagation procedure is controversal
  \citep{goodfellow2016deep}.
  \item In sum,
  the MDP account may serve as a concrete conceptualization of the PC account
  in cognitive neuroscience. MPDs have the advantage of exposing an explicit
  mechanism of the horizon of future considerations or
  how the internal model of the world gets updated,
  and collapses sensory input processing and action output preparation.
  %
  As such, our MDP account of DMN function can parsimoneously explain
  prediction coding hypothesis.
\end{itemize}

@@

\begin{itemize}
  \item Both the semantic hypothesis and MDPs expose mechanisms of
  how alternative decision trees could be mentally explored.
  \item In both semantic hypothesis and MDPs,
  there is no evidence to indicate that predictions of various
  levels of complexity, abstraction, timescale and purpose
  use mechanisms that are qualitatively different. This concurs with
  DMN activity increases across time, space, and content domains
  demonstrated in many neuroimaging studies. The semantic hypothesis
  and RL account provide explanations why hippocampus lesion does
  not only impair retrieving memories, but also hypothetical and future
  thinking \citep{hassabis2007patients}.
  \item The notion of semantic or knowledge association is
  incorporated into the MDP as the Markov property,
  that is, the current state directly results from the
  agent's history of states and actions. The learned
  value matrix and action transition schedules drive
  stimulus processing and action choice in the present.
  \item Mental scene construction has been proposed by some authors
  \citep{buckner2007self} to imply a distinction between
    environmental perception and internally generated mind-wandering\footnote{
      \textbf{XXX (danilo) This is more than just a footnote. Make this part of the paper
      and factorize the lot. Also you probably need a citation here!!!}
  ``A computational model of how such a process might be structured
  is far from being defined, but it will probably require a form of
  regulation by which perception of the current world is suppressed
  while simulation of possible alternatives are constructed,
  followed by a return to perception of the present.''\citep{buckner2007self}}.
  The MDPs naturally integrate the former egocentric
  (more related to current action, state and reward) and the later
  allocentric (more related to past and future actions, states, and rewards)
  angles on the world in a same optimization problem.
  \item The semantic account of DMN function does not offer
  a mechanistic explanation how explicit world knowledge and semantic analogies thereof
  lead to prediction of future actions and states.
  \item In contrast to existing accounts on semantics and
  mental scene construction, the random and creative aspects of DMN function
  are explained in MDPs by the advantages of stochastic optimization.
  \item The semantic hypothesis does not explain why memory recall
  for scene construction in humans is typically fragmentary and noisy
  instead of accurate and reliable.
  Yet, the MDP framework provides an algorithmic explanation in that
  stochasticity of the parameter space explored
  by the Monte Carlo solvers achieves better fine-tuning of the
  action policies and estimation of expected reward outcomes.
  That is, the purposeful stochasticity of policy and value estimation
  in MDPs provides a candidate explanation for why humans
  have evolved imperfect noisy memories
  as the more advantageous adaptation.
  \item While both semantic and MDP account propose memory-based internally
  generated information for probabilistic mental models of action outcomes,
  only MDPs render explicit the grounds on which the final action is
  eventually chosen (i.e., the estimated cumulative reward).
  \item In sum, episodic scene construction according to the semantic
  hypothesis is lacking an explicit time and incentive model;
  neither does it explain how randomness of human mental experience
  can be beneficial.
\end{itemize}

@@

\textbf{XXX (danilo) Your focus in this subsection ought to be on the SC itself not POMPDs. -> agree I will fix it later.}
 \begin{itemize}
    \item Processing social information has been proposed to underlie the
    physiological baseline of human brain function
    \citep{schilbach2008minds}. This was later challenged by observing
    analogues of the DMN in monkeys \citep{mantini2011default}
    and rats \citep{lu2012rat}, with
    supposedly less advanced social capacities.
    The MDPs parsimoneously explain the dominance of social content in
    human mental scenes by their extremely high relevance to humans
    reflected by high values for social information in the value matrix
    \citep{baker2009action}.
    \item How can a same neurobiological circuit be equally important
    for baseline house-keeping functions and specific task performance?
    Our computational account can explain why the DMN is implicated
    in both a goal-directed task and an idlying rest cognitive set,
    if environmental relevance is processed, manipulated, and retained
    also as a baseline function of human mental activity.
    In tasks, the policy and value matrices are updated to optimize short-term action,
    whereas, at rest, these parameter updates may
    improve especially mid- and long-term action.
    \item The sentinel account does not provide a formal mechanisms of
    how attention and memory resources are exactly reallocated when
    encountering a salient environmental stimulus. \textbf{XXX remove the next sentence. Danilo: can you help me to correct the sentence instead.
    There appears to be a cool link between cognition and formal theory
    here.}
    In contrast,
    the Bellman recursion is weighted at each step by
    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
    are recursively explored by the agent and which decision trees are neglected.
    This directly implies allocation of attention and computational load.
    \item In sum,
    MDPs imply a ``radar'' function of monitoring the environment for salient information
    in general and explain social thinking as a physiological brain baseline
    as an important special case, without excluding other topics
    contemplated in continuous mental activity.
 \end{itemize}
